---
title: "Objective 1"
author: "Austin"
date: "5/6/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}

library(fivethirtyeight)
library(tidyverse)
library(lubridate)
library(RColorBrewer)
library(formattable)

data("classic_rock_raw_data")
data("classic_rock_song_list")

```


### **Objective 1**
#### The first objective is to analyze the most popular songs and artists by decade. This objective shows us not what was most popular at the time of release but what songs/artists have held on to their popularity and are played often today.  

**Question 1: What are the most played songs for each decade?**  
The first step to answering this question is to create intervals of our release_year variable using the cut function.
```{r}

rock <- classic_rock_song_list

# Cutting our Release dates into intervals to show decades of Rock.
rock$release_year <- cut(
    rock$release_year,
    breaks = c(1950, 1960, 1970, 1980, 1990, 2000, 2010, 2020),
    dig.lab = 4
  ) 

```

Then we are going to filter our results by has_year to make sure the song has a release year, then filter by our decade, group by song, and finally tally the playcounts and display the top 9 songs.

```{r}

rock %>%
  filter(has_year == TRUE) %>%
  filter(release_year == "(1950,1960]") %>%
  group_by(song) %>%
  tally(playcount) %>%
  arrange_( ~ desc(n)) %>%
  top_n(9) 

```

Next we will be adding the ggplot component to display our findings in a barchart. We are using a barchart with flipped coordinates to properly display song titles and an Iowa State color scheme to match our presentation. 

```{r}
rock %>%
  filter(has_year == TRUE) %>%
  filter(release_year == "(1950,1960]") %>%
  group_by(artist) %>%
  tally(playcount) %>%
  arrange_( ~ desc(n)) %>%
  top_n(9) %>%
  ggplot(aes(
    x = reorder(artist, n),
    y = n,
    fill = factor(artist)
  )) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() +
  labs(y = "Playcount", x = "Artist",  title = "1950s Playcount by Artist") +
  theme_minimal() +
  scale_fill_brewer(palette = "YlOrRd")

```

All of song graphs below

```{r}
rock %>%
  filter(has_year == TRUE) %>%
  filter(release_year == "(1950,1960]") %>%
  group_by(artist) %>%
  tally(playcount) %>%
  arrange_( ~ desc(n)) %>%
  top_n(9) %>%
  ggplot(aes(
    x = reorder(artist, n),
    y = n,
    fill = factor(artist)
  )) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() +
  labs(y = "Playcount", x = "Artist",  title = "1950s Playcount by Artist") +
  theme_minimal() +
  scale_fill_brewer(palette = "YlOrRd")


rock  %>%
  filter(has_year == TRUE) %>%
  filter(release_year == "(1960,1970]") %>%
  group_by(artist) %>%
  tally(playcount) %>%
  arrange_( ~ desc(n)) %>%
  top_n(9) %>%
  ggplot(aes(
    x = reorder(artist, n),
    y = n,
    fill = factor(artist)
  )) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() +
  labs(y = "Playcount", x = "Artist",  title = "1960s Playcount by Artist") +
  theme_minimal() +
  scale_fill_brewer(palette = "YlOrRd")

rock  %>%
  filter(has_year == TRUE) %>%
  filter(release_year == "(1970,1980]") %>%
  group_by(artist) %>%
  tally(playcount) %>%
  arrange_( ~ desc(n)) %>%
  top_n(9) %>%
  ggplot(aes(
    x = reorder(artist, n),
    y = n,
    fill = factor(artist)
  )) +
  
  
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() +
  labs(y = "Playcount", x = "Artist",  title = "1970s Playcount by Artist") +
  theme_minimal() +
  scale_fill_brewer(palette = "YlOrRd")


rock  %>%
  filter(has_year == TRUE) %>%
  filter(release_year == "(1980,1990]") %>%
  group_by(artist) %>%
  tally(playcount) %>%
  arrange_( ~ desc(n)) %>%
  top_n(9) %>%
  ggplot(aes(
    x = reorder(artist, n),
    y = n,
    fill = factor(artist)
  )) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() +
  labs(y = "Playcount", x = "Artist",  title = "1980s Playcount by Artist") +
  theme_minimal() +
  scale_fill_brewer(palette = "YlOrRd")


rock  %>%
  filter(has_year == TRUE) %>%
  filter(release_year == "(1990,2000]") %>%
  group_by(artist) %>%
  tally(playcount) %>%
  arrange_( ~ desc(n)) %>%
  top_n(9) %>%
  ggplot(aes(
    x = reorder(artist, n),
    y = n,
    fill = factor(artist)
  )) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() +
  labs(y = "Playcount", x = "Artist",  title = "1990s Playcount by Artist") +
  theme_minimal() +
  scale_fill_brewer(palette = "YlOrRd")



rock %>%
  filter(has_year == TRUE) %>%
  filter(release_year == "(2000,2010]") %>%
  group_by(artist) %>%
  tally(playcount) %>%
  arrange_( ~ desc(n)) %>%
  top_n(7) %>%
  ggplot(aes(
    x = reorder(artist, n),
    y = n,
    fill = factor(artist)
  )) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() +
  labs(y = "Playcount", x = "Artist",  title = "2000s Playcount by Artist") +
  theme_minimal() +
  scale_fill_brewer(palette = "YlOrRd")


rock %>%
  filter(has_year == TRUE) %>%
  filter(release_year == "(2010,2020]") %>%
  group_by(artist) %>%
  tally(playcount) %>%
  arrange_( ~ desc(n)) %>%
  top_n(9) %>%
  ggplot(aes(
    x = reorder(artist, n),
    y = n,
    fill = factor(artist)
  )) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() +
  labs(y = "Playcount", x = "Artist",  title = "2010s Playcount by Artist") +
  theme_minimal() +
  scale_fill_brewer(palette = "YlOrRd")


```

**Question 2: What are the most played artists for each decade?**
We will be doing a similar set of steps to create graphs based on artists. Starting from our cut decades, filtering by has_year and release_year, grouping by artist, tallying playcount, displaying top 9 artists

```{r}
rock %>%
  filter(has_year == TRUE) %>%
  filter(release_year == "(1950,1960]") %>%
  group_by(artist) %>%
  tally(playcount) %>%
  arrange_( ~ desc(n)) %>%
  top_n(9) 

```

Once again adding the ggplot component to display our findings in a barchart. We are using a barchart with flipped coordinates to properly display artist names and an Iowa State color scheme to match our presentation. 
```{r}
rock %>%
  filter(has_year == TRUE) %>%
  filter(release_year == "(1950,1960]") %>%
  group_by(artist) %>%
  tally(playcount) %>%
  arrange_( ~ desc(n)) %>%
  top_n(9) %>%
  ggplot(aes(
    x = reorder(artist, n),
    y = n,
    fill = factor(artist)
  )) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() +
  labs(y = "Playcount", x = "Artist",  title = "1950s Playcount by Artist") +
  theme_minimal() +
  scale_fill_brewer(palette = "YlOrRd")

```

All of our artist graphs below

```{r}
rock %>%
  filter(has_year == TRUE) %>%
  filter(release_year == "(1950,1960]") %>%
  group_by(song) %>%
  tally(playcount) %>%
  arrange_(~ desc(n)) %>%
  top_n(8) %>%
  ggplot(aes(
    x = reorder(song, n),
    y = n,
    fill = factor(song)
  )) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  
  coord_flip() +
  labs(y = "Playcount", x = "Song",  title = "1950s Playcount by Song") +
  
  theme_minimal() +
  
  scale_fill_brewer(palette = "YlOrRd")

rock %>%
  filter(has_year == TRUE) %>%
  filter(release_year == "(1960,1970]") %>%
  group_by(song) %>%
  tally(playcount) %>%
  arrange_(~ desc(n)) %>%
  top_n(8) %>%
  ggplot(aes(
    x = reorder(song, n),
    y = n,
    fill = factor(song)
  )) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() +
  labs(y = "Playcount", x = "Song",  title = "1960s Playcount by Song") +
  theme_minimal() +
  scale_fill_brewer(palette = "YlOrRd")

rock %>%
  filter(has_year == TRUE) %>%
  filter(release_year == "(1970,1980]") %>%
  group_by(song) %>%
  tally(playcount) %>%
  arrange_(~ desc(n)) %>%
  top_n(8) %>%
  ggplot(aes(
    x = reorder(song, n),
    y = n,
    fill = factor(song)
  )) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() +
  labs(y = "Playcount", x = "Song",  title = "1970s Playcount by Song") +
  theme_minimal() +
  scale_fill_brewer(palette = "YlOrRd")

rock  %>%
  filter(has_year == TRUE) %>%
  filter(release_year == "(1980,1990]") %>%
  group_by(song) %>%
  tally(playcount) %>%
  arrange_(~ desc(n)) %>%
  top_n(8) %>%
  ggplot(aes(
    x = reorder(song, n),
    y = n,
    fill = factor(song)
  )) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() +
  labs(y = "Playcount", x = "Song",  title = "1980s Playcount by Song") +
  theme_minimal() +
  scale_fill_brewer(palette = "YlOrRd")

rock  %>%
  filter(has_year == TRUE) %>%
  filter(release_year == "(1990,2000]") %>%
  group_by(song) %>%
  tally(playcount) %>%
  arrange_(~ desc(n)) %>%
  top_n(8) %>%
  ggplot(aes(
    x = reorder(song, n),
    y = n,
    fill = factor(song)
  )) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() +
  labs(y = "Playcount", x = "Song",  title = "1990s Playcount by Song") +
  theme_minimal() +
  scale_fill_brewer(palette = "YlOrRd")

rock  %>%
  filter(has_year == TRUE) %>%
  filter(release_year == "(2000,2010]") %>%
  group_by(song) %>%
  tally(playcount) %>%
  arrange_(~ desc(n)) %>%
  top_n(8) %>%
  ggplot(aes(
    x = reorder(song, n),
    y = n,
    fill = factor(song)
  )) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() +
  labs(y = "Playcount", x = "Song",  title = "2000s Playcount by Song") +
  theme_minimal() +
  scale_fill_brewer(palette = "YlOrRd")

rock %>%
  filter(has_year == TRUE) %>%
  filter(release_year == "(2010,2020]") %>%
  group_by(song) %>%
  tally(playcount) %>%
  arrange_(~ desc(n)) %>%
  top_n(8) %>%
  ggplot(aes(
    x = reorder(song, n),
    y = n,
    fill = factor(song)
  )) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() +
  labs(y = "Playcount", x = "Song",  title = "2010s Playcount by Song") +
  theme_minimal() +
  scale_fill_brewer(palette = "YlOrRd")


```


**Question 3: Which decade had the most plays?**
To answer our last question of the most played decade, instead of grouping by artist or song we will be grouping by decade, tallying playcounts and displaying the results in ggplot with flipped coordinates to properly display decades and an Iowa State color scheme to match our presentation.

```{r}
rock %>%
  filter(has_year == TRUE) %>%
  filter(release_year == "(1950,1960]") %>%
  group_by(artist) %>%
  tally(playcount) %>%
  arrange_( ~ desc(n)) %>%
  top_n(9) %>%
  ggplot(aes(
    x = reorder(artist, n),
    y = n,
    fill = factor(artist)
  )) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() +
  labs(y = "Playcount", x = "Artist",  title = "1950s Playcount by Artist") +
  theme_minimal() +
  scale_fill_brewer(palette = "YlOrRd")
```

<br>
